---
title: "Brewery Case Study"
author: "Julia Layne"
date: "2/18/2020"
output: html_document
---

## Intro
The introduction needs to be written as if you are presenting the work to the CEO and CFO of Budweiser (your client) and that they only have had one class in statistics.  If it sounds like a student presentation, that is not acceptable.  You may assume that the CEO and CFO gave you the data and gave you the directive to report any interesting finding that you may uncover through your analysis.

## Code Purpose
b.	Briefly explain the purpose of the code. The explanations should appear as a sentence or two before or after the code chunk. Even though you will not be hiding the code chunks (so that I can see the code), you need to assume that the client can’t see them.

## Questions
c.	Use R to code answers concerning the seven questions below.

### Import
#### Files
```{r}
beer_raw = read.csv(file.choose(),header = TRUE)
brew_raw = read.csv(file.choose(),header = TRUE)
```
#### Libaries
```{r}
#install.packages('usmap')
library(dplyr)
library(usmap)
library(ggplot2)
```
### Breweries By State
1.   How many breweries are present in each state?
```{r}
brew_count_by_state <- brew_raw %>% group_by(State) %>% tally()
brew_count_by_state$state <- trimws(as.character(brew_count_by_state$State))
brew_count_by_state$fips <- fips(brew_count_by_state$state)
attach(brew_count_by_state)
brew_count_fips <- brew_count_by_state[order(fips),] 
detach(brew_count_by_state)

brew_count_fips

plot_usmap(data = brew_count_fips,  values = "n", color = "blue") + 
    labs(title = "Breweries by State", subtitle = "Count of Breweries per state") + 
  scale_fill_continuous(low = "white", high = "blue", name = "Breweries by State", label = scales::comma) + theme(legend.position = "right")
```

### Merged Data
2.   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)
```{r}

beer_raw
brew_raw$Brewery_id = brew_raw$Brew_ID


beer_by_brew <- merge(beer_raw,brew_raw,by="Brewery_id")
head(beer_by_brew) 
tail(beer_by_brew)

```

### Missing Data
3.   Address the missing values in each column.
```{r}
#summary(beer_by_brew)
#ABV, IBU Just omitting for now, come back
beer_by_brew_clean <- na.omit(beer_by_brew)
summary(beer_by_brew_clean)
brew_count_by_state <- beer_by_brew_clean %>% group_by(State) %>% tally()
brew_count_by_state$state <- trimws(as.character(brew_count_by_state$State))
brew_count_by_state$fips <- fips(brew_count_by_state$state)
attach(brew_count_by_state)
brew_count_fips <- brew_count_by_state[order(fips),] 
detach(brew_count_by_state)
```

### ABV and IBU by State
4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}

#how could I bucket these for better histogram grouping?

medianABVs <- as.data.frame(beer_by_brew_clean %>% group_by(State) %>% summarise( MedianABV=median(ABV* 100), MeanABV=mean(ABV* 100), MaxABV=max(ABV* 100), MinABV=min(ABV* 100), StdABV=sd(ABV* 100)))
medianIBUs <- as.data.frame(beer_by_brew_clean %>% group_by(State) %>% summarise( MedianIBU=median(IBU), MeanIBU=mean(IBU), MaxIBU=max(IBU), MinIBU=min(IBU), StdIBU=sd(IBU))) 

medianABVIBU <- merge(medianABVs,medianIBUs,by="State")
medianABVIBU$state <- trimws(as.character(medianABVIBU$State))
medianABVIBU$fips <- fips(medianABVIBU$state)
attach(medianABVIBU)
medianABVIBU_fips <- medianABVIBU[order(fips),] 
detach(medianABVIBU)

medianABVIBU_fips

plot_usmap(data = medianABVIBU_fips,  values = "MedianIBU", color = "blue") + 
    labs(title = "Median IBU by State", subtitle = "International Bitterness Units") + 
  scale_fill_continuous(low = "white", high = "blue", name = "IBU", label = scales::comma) + theme(legend.position = "right")

plot_usmap(data = medianABVIBU_fips,  values = "MedianABV", color = "red") + 
    labs(title = "Median ABV by State", subtitle = "Alcohol by Volume") + 
  scale_fill_continuous(low = "white", high = "red", name = "ABV", label = scales::comma) + theme(legend.position = "right")

plot_usmap(include= c("OK", "TX", "CO", "NM"), data = medianABVIBU_fips,  values = "MedianABV", color = "red") + 
    labs(title = "Median ABV by State", subtitle = "Alcohol by Volume") + 
  scale_fill_continuous(low = "white", high = "red", name = "ABV", label = scales::comma) + theme(legend.position = "right")

plot_usmap(include= c("MN", "IA", "MO", "AR", "LA"), data = medianABVIBU_fips,  values = "MedianABV", color = "red") + 
    labs(title = "Median ABV by State", subtitle = "Alcohol by Volume") + 
  scale_fill_continuous(low = "white", high = "red", name = "ABV", label = scales::comma) + theme(legend.position = "right")
```


### Highest ABV and IBU
5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
max_ABV <-medianABVIBU_fips[order(-medianABVIBU_fips$MaxABV),] 
head(max_ABV, 1)
max_IBU <- medianABVIBU_fips[order(-medianABVIBU_fips$MaxIBU),] 
head(max_IBU, 1)
max_ABV
includePrintABV = head(max_ABV, 5)$state
includePrintIBU = head(max_IBU, 5)$state
includePrintABV
includePrintIBU

plot_usmap(include= includePrintABV, data = medianABVIBU_fips,  values = "MaxABV", color = "red") + 
    labs(title = "Top 5 ABV States", subtitle = "Alcohol by Volume") + 
  scale_fill_continuous(low = "white", high = "red", name = "ABV", label = scales::comma) + theme(legend.position = "right")

plot_usmap(include= includePrintIBU, data = medianABVIBU_fips,  values = "MaxIBU", color = "red") + 
    labs(title = "Top 5 IBU States", subtitle = "International Bitterness Units") + 
  scale_fill_continuous(low = "white", high = "red", name = "IBU", label = scales::comma) + theme(legend.position = "right")
```

### Summary Statistics
6.   Comment on the summary statistics and distribution of the ABV variable.

### IBU and ABV Relationship
7.   Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

### IBU and ABV: IPA vs Ales
8.  Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually.

In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

9. Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 








